### 2023-02-20 paper-thesis/chapters/01-introduction.tex
Add example to background

Type checking programming languages is commonly done by writing a type checker in a general purpose language. This requires a lot of effort, and it is easy to make mistakes. In this thesis we will specifically focus on dependently typed languages, which differ from other languages because they allow types to be parameterized by values. This allows types to express properties of values that cannot be expressed in a simple type system, such as the length of a list or the well-formedness of a binary search tree. This expressiveness also makes dependent type systems more complicated to type check, since deciding equality of types requires normalization of the terms they are parameterized by.

>>>>

Type checking is commonly done by writing a type checker in a general purpose language. This requires a lot of effort, and it is easy to make mistakes. In this thesis we will specifically focus on dependently typed languages, which differ from other languages because they allow types to be parameterized by values~\cite{lambda_cube}. This allows types to express properties of values that cannot be expressed in a simple type system, such as the length of a list or the well-formedness of a binary search tree. This expressiveness also makes dependent type systems more complicated to type check, since deciding equality of types requires normalization of the terms they are parameterized by~\cite{Coquand_Huet_1988}.

### 2023-02-20 paper-thesis/chapters/02-background-spoofax.tex
Syntax highlighting
Chapter transitions
Add example to background

\paragraph{Syntax definitions}
The syntax of a language is defined through rules of the form \verb|A.C = a| with \verb|A| being a \emph{Non-Terminal}, \verb|C| being the name of a constructor and \verb|a| a list of symbols. To remove ambiguity, one can specify a relative priority for each constructor. For example, below is the syntax of a simple language:
\begin{lstlisting}
context-free sorts
	Expr
context-free syntax
	Expr.Add = Expr "+" Expr
	Expr.Mul = Expr "*" Expr
	Expr.X = "x"
context-free priorities
	Expr.X > Expr.Mul > Expr.Add
\end{lstlisting}

@@

>>>>

\paragraph{Syntax definitions}
The syntax of a language is defined through rules of the form \verb|A.C = a| with \verb|A| being a \emph{Non-Terminal}, \verb|C| being the name of a constructor and \verb|a| a list of symbols. To remove ambiguity, one can specify a relative priority for each constructor. For example, below is the syntax of a simple language:

\lstset{language=SDF3}
\begin{lstlisting}
context-free sorts
	Expr
context-free syntax
	Expr.Add = Expr "+" Expr
	Expr.Mul = Expr "*" Expr
	Expr.X = "x"
context-free priorities
	Expr.X > Expr.Mul > Expr.Add
\end{lstlisting}
\lstset{language=base}

@@

\subsection*{Example}
To get a better understanding of scope graphs, we consider an example in a non-dependent language. 
\lstset{language=java}
\begin{lstlisting}
	int x = 5;
	int y = x + 3;
	return x + y;
\end{lstlisting}
Remember that you can think of scopes as a type-checking environment. During type checking, the following steps happen:
\begin{enumerate}
	\item \verb|5| is type-checked in the root scope \verb|s0|. 
	\item The definition of \verb|x| creates a new scope \verb|s1| that contains the type of \verb|x|, which is used to type-check the rest of the program.
	\item \verb|x + 3| is type-checked in the new scope \verb|s1|, the \verb|x| is resolved to the declaration in \verb|s1|.
	\item The definition of \verb|y| creates a new scope \verb|s2| that contains the type of \verb|y|, which is used to type-check the rest of the program.
	\item \verb|x + y| is type-checked in the new scope \verb|s2|, to resolve \verb|x| we need to follow a \verb|P| edge.
\end{enumerate}

\lstset{language=}
\begin{tikzpicture}[scopegraph, node distance = 2.5em and 2.5em]
	\node[scope] (s0) {$s0$};

	\node[scope, right of = s0, xshift=5em] (s1) {$s1$};
	\draw (s1) edge[lbl={P}] (s0);
	\node[type, above of = s1] (s1l) {$\texttt{x : Int()}$};
	\draw (s1) edge[type] (s1l);

	\node[scope, right of = s1, xshift=6em] (s2) {$s2$};
	\draw (s2) edge[lbl={P}] (s1);
	\node[type, above of = s2] (s2l) {$\texttt{y : Int()}$};
	\draw (s2) edge[type] (s2l);
\end{tikzpicture}
